---
title: "Camalytics Regression Analysis"
author: "Cam Smithers"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Camalytics

## Overview

## Loading Packages & Exponential Variable

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(foreign)
library(nnet)
library(readr)
library(patchwork)
#library(plotly)
```

## Loading in All Data

```{r warning=FALSE}
#Running script that contains all data (teams and players)
source("/Users/camsmithers/Desktop/NBA Project/Main/Data/DataScript.R")

#Joining all team data into one data frame
all_team_data <- team_traditional %>%
    left_join(team_advanced, by = c("team", 'date')) %>% 
    left_join(team_misc, by = c("team", "date")) %>%
    left_join(team_scoring, by = c("team", "date")) %>%
    left_join(team_outcome, by = c("team", "date"))
```

## Data Cleaning 1: General Cleaning

```{r}
all_team_data <- all_team_data %>%
    dplyr::select(-opponent, -opponent.y, -opponent.x.x, -opponent.y.y, -home,
                  -home.y, -home.x.x,
           -home.y.y, -blk.y, -pf.y) %>% #Removing duplicated columns
    rename("opponent" = "opponent.x", "blk" = "blk.x", "pf" = "pf.x",
           "home" = "home.x") %>% #Wanted to keep specific duplicated to be used as the main
    dplyr::select(team, home, everything()) #Re-ordering variables for better visualization
team_temporary_df <- all_team_data %>% #Temporary data frame used add opponents' points scored
    dplyr::select(team, opponent, date,
                  #Traditional Statistics
                  pts, fgpct,
                  #Advanced Statistics
                  threefgpct, efgpct, tspct) %>%
    rename("team_1" = "team", "team_2" = "opponent",
           "opponent_pts" = "pts", "opponent_fgpct" = "fgpct",
           "opponent_threefgpct" = "threefgpct", "opponent_efgpct" = "efgpct",
           "opponent_tspct" = "tspct") %>%
    dplyr::select(-team_1)

all_team_data <- all_team_data %>% #Joining opponent points to original data frame
    left_join(team_temporary_df, by = c("team" = "team_2", "date"))

all_team_data <- all_team_data %>% #Re-ordering for better visualization
    dplyr::select(date, home, team, pts, opponent, opponent_pts, everything()) %>%
    rename_with(~ paste0("team_", .)) #Adding _team to each observation for easier recognition of variable for later cleaning, model building, and joins
```

## Data Cleaning 2: Creating Variables for Outcomes

```{r}
all_team_data <- all_team_data %>%
    #Variables for Later Mutation
    mutate(team_win = ifelse(team_plusminus > 0, 1, 0),
           team_loss = ifelse(team_plusminus < 0, 1, 0),
           team_pointdiff = abs(team_plusminus),
           game_points = team_pts + team_opponent_pts) %>%
    
    #Outcome 1: Margin of Victory
    mutate(team_game_outcome = case_when(
        team_plusminus > 0 & team_plusminus <= 10 ~ 1,
        team_plusminus < 0 & team_plusminus >= -10 ~ 2,
        team_plusminus > 10 ~ 3,
        team_plusminus < -10 ~ 4)) %>%
    
    #Outcome 2: Game Outcome 
    mutate(team_game_outcome_2 = ifelse(team_plusminus > 0, 1, 0)) %>%
    
    #Outcome 3: Game Upset
        #See Upset Model for mutations
    
    #Outcome 4: Covering Point Spread
    mutate(team_game_outcome_4 = case_when(
        team_pointdiff > team_spread & team_favored == 1 ~ 1,
        team_pointdiff < team_spread & team_favored == 1 ~ 0,
        team_pointdiff == team_spread & team_favored == 1 ~ 0,
        team_pointdiff < team_spread & team_favored == 0 ~ 1, 
        team_pointdiff > team_spread & team_favored == 0 ~ 0,
        team_pointdiff == team_spread & team_favored == 0 ~ 1)) %>%
    #Outcome 5: Covering Game Total
    mutate(team_game_outcome_5 = ifelse(game_points > team_gametotal, 1, 0))
```

## Data Cleaning 3: Team Averages

```{r}
all_team_average <- all_team_data %>%
    group_by(team_team) %>%
    mutate(
        #Offensive Statistics
        team_cummean_pts = round(cummean(team_pts), 1),
        team_cummean_offrating = round(cummean(team_offrating), 1),
        team_cummean_fgpct = round(cummean(team_fgpct), 1),
        team_cummean_efgpct = round(cummean(team_efgpct), 1),
        team_cummean_tspct = round(cummean(team_tspct), 1),
        
        #Defensive Statistics
        team_cummean_pts_allowed = round(cummean(team_opponent_pts), 1),
        team_cummean_defrating = round(cummean(team_defrating), 1),
        team_cummean_fgpct_allowed = round(cummean(team_opponent_fgpct), 1),
        team_cummean_efgpct_allowed = round(cummean(team_opponent_fgpct), 1),
        team_cummean_tspct_allowed = round(cummean(team_opponent_tspct), 1),
        
        #Other Statistics
        team_cummean_reb = round(cummean(team_reb), 1),
        team_cummean_tov = round(cummean(team_tov), 1),
        team_cummean_pts_off_tov = round(cummean(team_pts_off_tov), 1),
        team_cummean_netrating = round(cummean(team_netrating), 1)) %>%
    fill(
        #Offensive Statistics
        team_cummean_pts, team_cummean_offrating, team_cummean_fgpct,
        team_cummean_efgpct, team_cummean_tspct,
        #Defensive Statistics
        team_cummean_pts_allowed, team_cummean_defrating, team_cummean_fgpct_allowed,
        team_cummean_efgpct_allowed, team_cummean_tspct_allowed,
        #Other Statistics
        team_cummean_reb, team_cummean_tov, team_cummean_netrating, 
        team_cummean_pts_off_tov, .direction = "down") %>%
    dplyr::select(
        team_team, team_date,
        #Offensive Statistics
        team_cummean_pts, team_cummean_offrating, team_cummean_fgpct,
        team_cummean_efgpct, team_cummean_tspct,
        #Defensive Statistics
        team_cummean_pts_allowed, team_cummean_defrating, team_cummean_fgpct_allowed,
        team_cummean_efgpct_allowed, team_cummean_tspct_allowed,
        #Other Statistics
        team_cummean_reb, team_cummean_tov, team_cummean_netrating, 
        team_cummean_pts_off_tov)

all_opponent_average <- all_team_data %>%
    group_by(team_team) %>%
    mutate(
        #Offensive Statistics
        opponent_cummean_pts = round(cummean(team_pts), 1),
        opponent_cummean_offrating = round(cummean(team_offrating), 1),
        opponent_cummean_fgpct = round(cummean(team_fgpct), 1),
        opponent_cummean_efgpct = round(cummean(team_efgpct), 1),
        opponent_cummean_tspct = round(cummean(team_tspct), 1),
        
        #Defensive Statistics
        opponent_cummean_pts_allowed = round(cummean(team_opponent_pts), 1),
        opponent_cummean_defrating = round(cummean(team_defrating), 1),
        opponent_cummean_fgpct_allowed = round(cummean(team_opponent_fgpct), 1),
        opponent_cummean_efgpct_allowed = round(cummean(team_opponent_fgpct), 1),
        opponent_cummean_tspct_allowed = round(cummean(team_opponent_tspct), 1),
        
        #Other Statistics
        opponent_cummean_reb = round(cummean(team_reb), 1),
        opponent_cummean_tov = round(cummean(team_tov), 1),
        opponent_cummean_pts_off_tov = round(cummean(team_pts_off_tov), 1),
        opponent_cummean_netrating = round(cummean(team_netrating), 1)) %>%
    fill(
        #Offensive Statistics
        opponent_cummean_pts, opponent_cummean_offrating, opponent_cummean_fgpct,
        opponent_cummean_efgpct, opponent_cummean_tspct,
        #Defensive Statistics
        opponent_cummean_pts_allowed, opponent_cummean_defrating, opponent_cummean_fgpct_allowed,
        opponent_cummean_efgpct_allowed, opponent_cummean_tspct_allowed,
        #Other Statistics
        opponent_cummean_reb, opponent_cummean_tov, opponent_cummean_netrating, 
        opponent_cummean_pts_off_tov, .direction = "down") %>%
    dplyr::select(
        team_team, team_date,
        #Offensive Statistics
        opponent_cummean_pts, opponent_cummean_offrating, opponent_cummean_fgpct,
        opponent_cummean_efgpct, opponent_cummean_tspct,
        #Defensive Statistics
        opponent_cummean_pts_allowed, opponent_cummean_defrating, opponent_cummean_fgpct_allowed,
        opponent_cummean_efgpct_allowed, opponent_cummean_tspct_allowed,
        #Other Statistics
        opponent_cummean_reb, opponent_cummean_tov, opponent_cummean_netrating, 
        opponent_cummean_pts_off_tov)

all_league_average <- all_team_data %>%
    group_by(team_date) %>%
    summarize(
        #Major Statistical Categories
        league_avg_pts = round(mean(team_pts, na.rm = TRUE), 1),
        league_avg_pts_allowed = round(mean(team_opponent_pts, na.rm = TRUE), 1),
        league_avg_reb = round(mean(team_reb, na.rm = TRUE), 1),
        league_avg_ast = round(mean(team_ast, na.rm = TRUE), 1),
        league_avg_tov = round(mean(team_tov, na.rm = TRUE), 1),
        #Advanced Metrics
        league_avg_offrating = round(mean(team_offrating, na.rm = TRUE), 1),
        league_avg_defrating = round(mean(team_defrating, na.rm = TRUE), 1),
        league_avg_netrating = round(mean(team_netrating, na.rm = TRUE), 1)) %>%
    arrange(team_date) %>%
    mutate(
        league_cummean_pts = round(cummean(league_avg_pts), 1),
        league_cummean_pts_allowed = round(cummean(league_avg_pts_allowed), 1),
        league_cummean_reb = round(cummean(league_avg_reb), 1),
        league_cummean_ast = round(cummean(league_avg_ast), 1),
        league_cummean_tov = round(cummean(league_avg_tov), 1),
        #Advanced Metrics
        league_cummean_offrating = round(cummean(league_avg_offrating), 1),
        league_cummean_defrating = round(cummean(league_avg_defrating), 1),
        league_cummean_netrating = round(cummean(league_avg_netrating), 1)) %>%
    fill(league_cummean_pts, league_cummean_pts_allowed, league_cummean_reb, 
         league_cummean_ast, league_cummean_tov, 
         league_cummean_offrating, league_cummean_defrating, league_cummean_netrating, .direction = "down") %>%
    dplyr::select(-starts_with("league_avg_"))

all_team_data <- all_team_data %>%
    left_join(all_team_average, by = c("team_team" = "team_team", "team_date" = "team_date")) %>%
    left_join(all_opponent_average, by = c("team_opponent" = "team_team", "team_date" = "team_date")) %>%
    left_join(all_league_average, by = "team_date")
```

## Data Cleaning 4:

```{r}
all_team_data <- all_team_data %>%
    group_by(team_team) %>%
    mutate(
        #Cumulative Values
        team_cumsum_win = cumsum(ifelse(is.na(team_win), 0, team_win)),
        team_cumsum_loss = cumsum(ifelse(is.na(team_loss), 0, team_loss)),
        team_cumsum_game = team_cumsum_win + team_cumsum_loss,
        #Cumulative Rates
        team_cumrate_win = ifelse(team_cumsum_game == 0, NA, round(team_cumsum_win / team_cumsum_game, 3))) %>%
    ungroup()
```

------------------------------------------------------------------------

# Logistic Regression & Linear Regression

## Outcome 2: Game Outcome: Binary Logistic Regression

```{r}
winloss_model <- glm(team_game_outcome_2 ~ team_home
                 + team_cummean_offrating
                 + team_cummean_defrating
                 + opponent_cummean_offrating
                 + opponent_cummean_defrating
                 , data = all_team_data, family = "binomial")

winloss_model_summary <- summary(winloss_model)
winloss_model_summary
```

### TGO 2: Model Predictors Data Frame

```{r}
#Model Estimate to Main Data Frame
winloss_predictions <- fitted(winloss_model)
winloss_predicted_values <- predict(winloss_model, newdata = all_team_data, type = "response")
all_team_data <- cbind(all_team_data, winloss_predicted_values)

#Cleaning Up Probability
all_team_data <- all_team_data %>%
    mutate(winloss_predicted_values = round(100 * winloss_predicted_values, 1))
```

## Outcome 3: Game Outcome (w/ Previous Condition Met): Binary Logistic Regression

```{r}
all_team_data <- all_team_data %>%
    mutate(winloss_model_guess = ifelse(winloss_predicted_values > 50, 1, 0)) %>%
    #Outcome 3: Predicting Upset
    mutate(team_game_outcome_3 = case_when(
        winloss_model_guess == 1 & team_plusminus < 0 ~ 1,
        winloss_model_guess == 1 & team_plusminus > 0 ~ 0,
        TRUE ~ NA))

all_team_data <- all_team_data %>%
    dplyr::group_by(team_team) %>%
    mutate(
    team_cumsum_favored = cumsum(ifelse(is.na(winloss_model_guess), 0, winloss_model_guess)),
    team_cumsum_upset = cumsum(ifelse(is.na(team_game_outcome_3), 0, team_game_outcome_3)),
    
    team_cumrate_favored = ifelse(team_cumsum_game == 0, NA, round(team_cumsum_favored / team_cumsum_game, 3)),
    team_cumrate_upset = ifelse(team_cumsum_favored == 0, NA, round(team_cumsum_upset / team_cumsum_favored, 3))) %>%
    ungroup()

upset_model <- glm(team_game_outcome_3 ~ team_cumrate_upset
                   + opponent_cummean_offrating
                   + opponent_cummean_defrating
                   , data = all_team_data, family = "binomial")
upset_model_summary <- summary(upset_model)
upset_model_summary
```

### TGO 3: Model Predictors Data Frame

```{r}
#Model Estimate to Main Data Frame
upset_predictions <- fitted(upset_model)
upset_predicted_values <- predict(upset_model, newdata = all_team_data, type = "response")
all_team_data <- cbind(all_team_data, upset_predicted_values)

#Cleaning Up Probability
all_team_data <- all_team_data %>%
    mutate(upset_predicted_values = round(100 * upset_predicted_values, 1)) %>%
    mutate(upset_predicted_values = case_when(
        winloss_model_guess == 1 ~ upset_predicted_values,
        winloss_model_guess == 0 ~ NA))
```

## Outcome 5: Game Total Over/Under: Linear Regression

```{r}
gametotal_model <- lm(game_points ~ team_cummean_pts 
                      + opponent_cummean_pts
                      + team_cummean_pts_allowed
                      + opponent_cummean_pts_allowed, data = all_team_data)
gametotal_model_summary <- summary(gametotal_model)
gametotal_model_summary
```

### TGO 5: Model Predictors to Data Frame

```{r}
gametotal_predicted_values <- predict(gametotal_model, newdata = all_team_data)
all_team_data <- cbind(all_team_data, gametotal_predicted_values)

all_team_data <- all_team_data %>%
    mutate(gametotal_predicted_values = round(gametotal_predicted_values))
```

------------------------------------------------------------------------

# Post Modeling Cleaning

```{r}
all_team_data <- all_team_data %>%
    mutate(outcome_2_estimates_categories = case_when(
        winloss_predicted_values >= 65 & upset_predicted_values <= 25 ~ "Very Likely", #Good
        
        winloss_predicted_values >= 65 & upset_predicted_values > 25 & upset_predicted_values < 50 ~ "Likely", #Good
        
        winloss_predicted_values >= 50 & winloss_predicted_values < 65 & upset_predicted_values <= 25 ~ "Likely", #Good
        
        winloss_predicted_values >= 50 & winloss_predicted_values < 65 
        & upset_predicted_values > 25 & upset_predicted_values < 50 ~ "Neutral", 
        
        winloss_predicted_values >= 65 & winloss_predicted_values >= 75 
        & upset_predicted_values >= 50 & upset_predicted_values <= 75 ~ "Unlikely"
        
        winloss_predicted_values >= 50 & winloss_predicted_values < 65 
        & upset_predicted_values >= 50 ~ "Very Unlikely"
    ))
```


prediction_zones <- data.frame(
  xmin = c(50, 50, 65, 65, 65, 50),
  xmax = c(65, 65, 100, 75, 100, 65),
  ymin = c(25, 50, 0, 50, 25, 0),
  ymax = c(50, 100, 25, 75, 50, 25),
  fill = c("3. Neutral", "5. Very Unlikely", "1. Very Likely", "4. Unlikely", "2. Likely", "2. Likely"))

------------------------------------------------------------------------

# Model Analysis

## Mutations for Analysis

```{r}
ma_all_team_data <- all_team_data %>%
    mutate(my_predicted_win = ifelse(winloss_predicted_values > 50, 1, 0),
           my_model_correct = ifelse(my_predicted_win == team_game_outcome_2, 1, 0),
           diff_from_fanduel = ifelse(my_predicted_win != team_favored, 1, 0), 
           my_predicted_upset_1 = ifelse(upset_predicted_values > 50, 1, 0),
           my_predicted_upset_2 = ifelse(upset_predicted_values > winloss_predicted_values, 1, 0)) %>%
    mutate(fanduel_correct = ifelse(team_favored == team_game_outcome_2, 1, 0)) %>%
    filter(!is.na(team_fgm)) %>%
    filter(team_cumsum_game >= 10) %>%
    mutate(my_predicted_win_factor = factor(my_predicted_win,
                                            levels = c(0, 1),
                                            labels = c("exploss", "expwin")),
           team_game_outcome_2_factor = factor(team_game_outcome_2,
                                               levels = c(0, 1),
                                               labels = c("loss", "win")),
           team_favored_factor = factor(team_favored,
                                        levels = c(0, 1),
                                        labels = c("exploss", "expwin")),
           my_predicted_upset_1_factor = factor(my_predicted_upset_1,
                                                levels = c(0, 1),
                                                labels = c("no exp upset", "exp upset")),
           my_predicted_upset_2_factor = factor(my_predicted_upset_2,
                                                levels = c(0, 1),
                                                labels = c("no exp upset", "exp upset")),
           team_game_outcome_3_factor = factor(team_game_outcome_3, 
                                               levels = c(0, 1),
                                               labels = c("no upset", "upset")))
```

## Model 2: Game Outcome

```{r}
cross_table_1 <- table(ma_all_team_data$my_predicted_upset_1_factor, ma_all_team_data$team_game_outcome_3_factor)
print(cross_table_1)
addmargins(cross_table_1)
prop.table(cross_table_1)

cross_table_2 <- table(ma_all_team_data$my_predicted_upset_2_factor, ma_all_team_data$team_game_outcome_3_factor)
print(cross_table_2)
addmargins(cross_table_2)
prop.table(cross_table_2)
```

## Model 3: Game Upset

------------------------------------------------------------------------

# Model Visualizations

## Mutations for Visualizations

```{r}
mv_all_team_data <- all_team_data %>%
    mutate(my_predicted_win = ifelse(winloss_predicted_values > 50, 1, 0),
           my_model_correct = ifelse(my_predicted_win == team_game_outcome_2, 1, 0),
           diff_from_fanduel = ifelse(my_predicted_win != team_favored, 1, 0), 
           my_predicted_upset = ifelse(upset_predicted_values > 50, 1, 0)) %>%
    mutate(fanduel_correct = ifelse(team_favored == team_game_outcome_2, 1, 0)) %>%
    filter(!is.na(team_fgm)) %>%
    filter(team_cumsum_game >= 10)

mv_one_team <- mv_all_team_data %>%
    dplyr::filter(team_team == "phx")
```

## Model 2: Game Outcome

```{r}
# Distribution of Percentages
winloss_plot_1 <- ggplot(mv_all_team_data, aes(x = winloss_predicted_values, fill = as.factor(team_game_outcome_2))) + 
    geom_histogram(color = "black", bins = 25) + 
    scale_x_continuous(breaks = seq(0, 100, by = 10))
winloss_plot_1

# Team Win Loss Probability
winloss_plot_2 <- ggplot(mv_all_team_data, aes(x = team_date, y = winloss_predicted_values, color = as.factor(team_game_outcome_2))) + 
    geom_point() +
    geom_path() + 
    facet_wrap(~team_team)
winloss_plot_2

# Scatter Plot
winloss_plot_3 <- ggplot(mv_all_team_data, aes(x = winloss_predicted_values, y = team_game_outcome_2, color = team_team)) + 
    geom_point() + 
    geom_smooth(method = "glm", method.args = list(family = "binomial"), se = TRUE, color = "black") + 
    facet_wrap(~team_team)
winloss_plot_3

prediction_zones <- data.frame(
  xmin = c(50, 50, 65, 65, 65, 50),
  xmax = c(65, 65, 100, 75, 100, 65),
  ymin = c(25, 50, 0, 50, 25, 0),
  ymax = c(50, 100, 25, 75, 50, 25),
  fill = c("3. Neutral", "5. Very Unlikely", "1. Very Likely", "4. Unlikely", "2. Likely", "2. Likely"))

winloss_plot_4 <- ggplot(mv_all_team_data, aes(x = winloss_predicted_values, y = upset_predicted_values, color = as.factor(team_game_outcome_2))) + 
    scale_x_continuous(limits = c(50, 100)) + 
    geom_hline(yintercept = 50, color = "black", linetype = "dashed") + 
    geom_vline(xintercept = 65, color = "white", linetype = "dashed") + 
    geom_rect(data = prediction_zones, 
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = fill), 
            alpha = 0.1, color = "black", inherit.aes = FALSE) + 
    scale_fill_manual(values = c(
    "3. Neutral" = "yellow",
    "5. Very Unlikely" = "red",
    "1. Very Likely" = "green",
    "4. Unlikely" = "orange",
    "2. Likely" = "blue",
    "2. Likely" = "blue")) + 
    geom_point() + 
    scale_color_discrete(labels = c("Loss", "Win")) + 
    labs(
    title = "Team Win & Upset Probability", 
    x = "Win Probability", 
    y = "Upset Probability", 
    color = "Game Outcome", 
    fill = "Probability Zones") + 
    theme_bw()
winloss_plot_4

ggsave(filename = "GameOutcome Splitting.png", plot = winloss_plot_4,
       path = "/Users/camsmithers/Desktop/NBA Project/Main/Visualizations", width = 25, height = 20, units = "cm")
```

```{r}
# Win vs Loss
winloss_plot_x <- ggplot(mv_all_team_data, aes(x = as.factor(team_game_outcome_2), fill = as.factor(my_predicted_win))) + 
    geom_bar(position = "dodge", color = "black") +
    geom_text(
        stat = "count",
        aes(label = after_stat(count)),
        position = position_dodge(width = 0.9),
        vjust = -0.3) #+ facet_wrap(~team_team)

winloss_plot_y <- ggplot(mv_all_team_data, aes(x = as.factor(team_game_outcome_2), fill = as.factor(team_favored))) + 
    geom_bar(position = "dodge", color = "black") +
    geom_text(
        stat = "count",
        aes(label = after_stat(count)),
        position = position_dodge(width = 0.9),
        vjust = -0.3) #+ facet_wrap(~team_team)

combined_plot <- winloss_plot_x + winloss_plot_y + 
    plot_layout(ncol = 2) + 
    plot_annotation(title = "Comparison of Predicted Wins and Team Favored",
                    theme = theme(plot.title = element_text(hjust = 0.5)))
combined_plot
```

## Model 3: Upset Prediction

## Model 4: Game Total

```{r}
gametotal_plot_1 <- ggplot(mv_all_team_data, aes(x = gametotal_predicted_values, y = game_points)) + 
    geom_point() +
    geom_smooth() +
    geom_jitter()
gametotal_plot_1
```


------------------------------------------------------------------------

# Machine Learning: Model Testing

## *Current Model: Margin of Victory*

## Necessary Packages

```{r eval=FALSE, include=FALSE}
library(rsample)
library(caret)
library(h2o)
library(modeldata)
library(vip)
library(ROCR)
```

## Needed Variables

```{r eval=FALSE, include=FALSE}
set.seed(1234)
ml_all_team_data <- all_team_data %>%
    dplyr::select(team_date, team_team, team_game_outcome, team_home, 
           team_moneyline)
ml_all_team_data <- na.omit(ml_all_team_data)
```

## Factoring Before Analysis

```{r eval=FALSE, include=FALSE}
ml_all_team_data <- ml_all_team_data %>%
    mutate(team_game_outcome = factor(team_game_outcome,
                                        levels = c(1, 2, 3, 4),
                                        labels = c("close_win", "close_loss",
                                                   "blowout_win", "blowout_loss")))
```

## Splitting Data & Creating Models

```{r eval=FALSE, include=FALSE}
#Splitting Data
ml_team_split <- initial_split(ml_all_team_data, prop = .60)
ml_team_train <- training(ml_team_split)
ml_team_test <- testing(ml_team_split)

#Models
ml_team_outcome_fd <- multinom(team_game_outcome ~ team_moneyline,
                               data = ml_team_train)

ml_team_outcome <- multinom(team_game_outcome ~ team_home 
                            + team_cummean_pts_win
                            + team_cummean_pts_allowed_win
                            + opponent_cummean_pts_loss
                            + opponent_cummean_pts_allowed_loss
                            + team_cummean_pts_loss
                            + team_cummean_pts_allowed_loss
                            + opponent_cummean_pts_win
                            + opponent_cummean_pts_allowed_win
                           , data = ml_team_train)

summary(ml_team_outcome_fd)
summary(ml_team_outcome)
```

## Cross Validation

```{r eval=FALSE, include=FALSE}
cv_ml_team_outcome_fd <- train(team_game_outcome ~ team_moneyline
                            , data = ml_team_train
                            , method = "multinom"
                            , trControl = trainControl(method = "cv", number = 5))

cv_ml_team_outcome <- train(team_game_outcome ~ team_home 
                            + team_cummean_pts_win
                            + team_cummean_pts_allowed_win
                            + opponent_cummean_pts_loss
                            + opponent_cummean_pts_allowed_loss
                            , data = ml_team_train
                           , method = "multinom"
                           , trControl = trainControl(method = "cv", number = 5))



summary(
    resamples(
        list(
            model1 = cv_ml_team_outcome_fd,
            model2 = cv_ml_team_outcome
        )
    )
)$statistics$Accuracy
```

```{r eval=FALSE, include=FALSE}
pred_class <- predict(cv_ml_team_outcome, ml_team_train)
confusionMatrix(
    data = pred_class,
    reference = relevel(ml_team_train$team_game_outcome, ref = 1)
)
```
