---
title: "Camalytics Regression Analysis"
author: "Cam Smithers"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Camalytics

## Overview

## Loading Packages & Exponential Variable

```{r}
library(ggplot2)
library(tidyverse)
library(foreign)
library(nnet)
library(readr)
```

## Loading in All Data

```{r warning=FALSE}
#Running script that contains all data (teams and players)
source("/Users/camsmithers/Desktop/NBA Project/Main/Data/DataScript.R")

#Joining all team data into one data frame
all_team_data <- team_traditional %>%
    left_join(team_advanced, by = c("team", 'date')) %>% 
    left_join(team_misc, by = c("team", "date")) %>%
    left_join(team_scoring, by = c("team", "date")) %>%
    left_join(team_outcome, by = c("team", "date"))
```

## Data Cleaning 1: General Cleaning

```{r}
all_team_data <- all_team_data %>%
    dplyr::select(-opponent, -opponent.y, -opponent.x.x, -opponent.y.y, -home,
                  -home.y, -home.x.x,
           -home.y.y, -blk.y, -pf.y) %>% #Removing duplicated columns
    rename("opponent" = "opponent.x", "blk" = "blk.x", "pf" = "pf.x",
           "home" = "home.x") %>% #Wanted to keep specific duplicated to be used as the main
    dplyr::select(team, home, everything()) #Re-ordering variables for better visualization
team_temporary_df <- all_team_data %>% #Temporary data frame used add opponents' points scored
    dplyr::select(team, opponent, date,
                  #Traditional Statistics
                  pts, fgpct,
                  #Advanced Statistics
                  threefgpct, efgpct, tspct) %>%
    rename("team_1" = "team", "team_2" = "opponent",
           "opponent_pts" = "pts", "opponent_fgpct" = "fgpct",
           "opponent_threefgpct" = "threefgpct", "opponent_efgpct" = "efgpct",
           "opponent_tspct" = "tspct") %>%
    dplyr::select(-team_1)

all_team_data <- all_team_data %>% #Joining opponent points to original data frame
    left_join(team_temporary_df, by = c("team" = "team_2", "date"))

all_team_data <- all_team_data %>% #Re-ordering for better visualization
    dplyr::select(date, home, team, pts, opponent, opponent_pts, everything()) %>%
    rename_with(~ paste0("team_", .)) #Adding _team to each observation for easier recognition of variable for later cleaning, model building, and joins
```

## Data Cleaning 2: Creating Variables for Outcomes

```{r}
all_team_data <- all_team_data %>%
    #Variables for Later Mutation
    mutate(team_win = ifelse(team_plusminus > 0, 1, 0),
           team_loss = ifelse(team_plusminus < 0, 1, 0),
           team_pointdiff = abs(team_plusminus),
           game_points = team_pts + team_opponent_pts) %>%
    
    #Outcome 1: Margin of Victory
    mutate(team_game_outcome_1 = case_when(
        team_plusminus %in% 1:10 ~ 0,
        team_plusminus >= 11 ~ 1,
        TRUE ~ NA)) %>%
    
    #Outcome 2: Game Outcome 
    mutate(team_game_outcome_2 = ifelse(team_plusminus > 0, 1, 0)) %>%
    
    #Outcome 3: Game Upset
        #See Upset Model for mutations
    
    #Outcome 4: Covering Point Spread
    mutate(team_game_outcome_4 = case_when(
        team_pointdiff > team_spread & team_favored == 1 ~ 1,
        team_pointdiff < team_spread & team_favored == 1 ~ 0,
        team_pointdiff == team_spread & team_favored == 1 ~ 0,
        team_pointdiff < team_spread & team_favored == 0 ~ 1, 
        team_pointdiff > team_spread & team_favored == 0 ~ 0,
        team_pointdiff == team_spread & team_favored == 0 ~ 1)) %>%
    #Outcome 5: Covering Game Total
    mutate(team_game_outcome_5 = ifelse(game_points > team_gametotal, 1, 0))
```

## Data Cleaning 3: Team Averages (General)

### Data Cleaning 3.1: Team Averages

```{r}
all_team_average <- all_team_data %>%
    group_by(team_team) %>%
    mutate(
        #Offensive Statistics
        team_cummean_pts = round(cummean(team_pts), 1),
        team_cummean_offrating = round(cummean(team_offrating), 1),
        team_cummean_fgpct = round(cummean(team_fgpct), 1),
        team_cummean_efgpct = round(cummean(team_efgpct), 1),
        team_cummean_tspct = round(cummean(team_tspct), 1),
        
        #Defensive Statistics
        team_cummean_pts_allowed = round(cummean(team_opponent_pts), 1),
        team_cummean_defrating = round(cummean(team_defrating), 1),
        team_cummean_fgpct_allowed = round(cummean(team_opponent_fgpct), 1),
        team_cummean_efgpct_allowed = round(cummean(team_opponent_fgpct), 1),
        team_cummean_tspct_allowed = round(cummean(team_opponent_tspct), 1),
        
        #Other Statistics
        team_cummean_reb = round(cummean(team_reb), 1),
        team_cummean_tov = round(cummean(team_tov), 1),
        team_cummean_pts_off_tov = round(cummean(team_pts_off_tov), 1),
        team_cummean_netrating = round(cummean(team_netrating), 1),
        team_cummean_plusminus = round(cummean(team_plusminus), 1)) %>%
    fill(
        #Offensive Statistics
        team_cummean_pts, team_cummean_offrating, team_cummean_fgpct,
        team_cummean_efgpct, team_cummean_tspct,
        #Defensive Statistics
        team_cummean_pts_allowed, team_cummean_defrating, team_cummean_fgpct_allowed,
        team_cummean_efgpct_allowed, team_cummean_tspct_allowed,
        #Other Statistics
        team_cummean_reb, team_cummean_tov, team_cummean_netrating, 
        team_cummean_pts_off_tov, team_cummean_plusminus, .direction = "down") %>%
    dplyr::select(
        team_team, team_date,
        #Offensive Statistics
        team_cummean_pts, team_cummean_offrating, team_cummean_fgpct,
        team_cummean_efgpct, team_cummean_tspct,
        #Defensive Statistics
        team_cummean_pts_allowed, team_cummean_defrating, team_cummean_fgpct_allowed,
        team_cummean_efgpct_allowed, team_cummean_tspct_allowed,
        #Other Statistics
        team_cummean_reb, team_cummean_tov, team_cummean_netrating, 
        team_cummean_pts_off_tov, team_cummean_plusminus)
```

### Data Cleaning 3.2: Opponent Averages

```{r}
all_opponent_average <- all_team_data %>%
    group_by(team_team) %>%
    mutate(
        #Offensive Statistics
        opponent_cummean_pts = round(cummean(team_pts), 1),
        opponent_cummean_offrating = round(cummean(team_offrating), 1),
        opponent_cummean_fgpct = round(cummean(team_fgpct), 1),
        opponent_cummean_efgpct = round(cummean(team_efgpct), 1),
        opponent_cummean_tspct = round(cummean(team_tspct), 1),
        
        #Defensive Statistics
        opponent_cummean_pts_allowed = round(cummean(team_opponent_pts), 1),
        opponent_cummean_defrating = round(cummean(team_defrating), 1),
        opponent_cummean_fgpct_allowed = round(cummean(team_opponent_fgpct), 1),
        opponent_cummean_efgpct_allowed = round(cummean(team_opponent_fgpct), 1),
        opponent_cummean_tspct_allowed = round(cummean(team_opponent_tspct), 1),
        
        #Other Statistics
        opponent_cummean_reb = round(cummean(team_reb), 1),
        opponent_cummean_tov = round(cummean(team_tov), 1),
        opponent_cummean_pts_off_tov = round(cummean(team_pts_off_tov), 1),
        opponent_cummean_netrating = round(cummean(team_netrating), 1),
        opponent_cummean_plusminus = round(cummean(team_plusminus), 1)) %>%
    fill(
        #Offensive Statistics
        opponent_cummean_pts, opponent_cummean_offrating, opponent_cummean_fgpct,
        opponent_cummean_efgpct, opponent_cummean_tspct,
        #Defensive Statistics
        opponent_cummean_pts_allowed, opponent_cummean_defrating, opponent_cummean_fgpct_allowed,
        opponent_cummean_efgpct_allowed, opponent_cummean_tspct_allowed,
        #Other Statistics
        opponent_cummean_reb, opponent_cummean_tov, opponent_cummean_netrating, 
        opponent_cummean_pts_off_tov, opponent_cummean_plusminus, .direction = "down") %>%
    dplyr::select(
        team_team, team_date,
        #Offensive Statistics
        opponent_cummean_pts, opponent_cummean_offrating, opponent_cummean_fgpct,
        opponent_cummean_efgpct, opponent_cummean_tspct,
        #Defensive Statistics
        opponent_cummean_pts_allowed, opponent_cummean_defrating, opponent_cummean_fgpct_allowed,
        opponent_cummean_efgpct_allowed, opponent_cummean_tspct_allowed,
        #Other Statistics
        opponent_cummean_reb, opponent_cummean_tov, opponent_cummean_netrating, 
        opponent_cummean_pts_off_tov, opponent_cummean_plusminus)
```

### Data Cleaning 3.3: League Averages

```{r}
all_league_average <- all_team_data %>%
    group_by(team_date) %>%
    summarize(
        #Major Statistical Categories
        league_avg_pts = round(mean(team_pts, na.rm = TRUE), 1),
        league_avg_pts_allowed = round(mean(team_opponent_pts, na.rm = TRUE), 1),
        league_avg_reb = round(mean(team_reb, na.rm = TRUE), 1),
        league_avg_ast = round(mean(team_ast, na.rm = TRUE), 1),
        league_avg_tov = round(mean(team_tov, na.rm = TRUE), 1),
        #Advanced Metrics
        league_avg_offrating = round(mean(team_offrating, na.rm = TRUE), 1),
        league_avg_defrating = round(mean(team_defrating, na.rm = TRUE), 1),
        league_avg_netrating = round(mean(team_netrating, na.rm = TRUE), 1)) %>%
    arrange(team_date) %>%
    mutate(
        league_cummean_pts = round(cummean(league_avg_pts), 1),
        league_cummean_pts_allowed = round(cummean(league_avg_pts_allowed), 1),
        league_cummean_reb = round(cummean(league_avg_reb), 1),
        league_cummean_ast = round(cummean(league_avg_ast), 1),
        league_cummean_tov = round(cummean(league_avg_tov), 1),
        #Advanced Metrics
        league_cummean_offrating = round(cummean(league_avg_offrating), 1),
        league_cummean_defrating = round(cummean(league_avg_defrating), 1),
        league_cummean_netrating = round(cummean(league_avg_netrating), 1)) %>%
    fill(league_cummean_pts, league_cummean_pts_allowed, league_cummean_reb, 
         league_cummean_ast, league_cummean_tov, 
         league_cummean_offrating, league_cummean_defrating, league_cummean_netrating, .direction = "down") %>%
    dplyr::select(-starts_with("league_avg_"))
```

### Data Cleaning 3.4: Joining Data

```{r}
all_team_data <- all_team_data %>%
    left_join(all_team_average, by = c("team_team" = "team_team", "team_date" = "team_date")) %>%
    left_join(all_opponent_average, by = c("team_opponent" = "team_team", "team_date" = "team_date")) %>%
    left_join(all_league_average, by = "team_date")
```

## Data Cleaning 4:

```{r}
all_team_data <- all_team_data %>%
    group_by(team_team) %>%
    mutate(
        #Cumulative Values
        team_cumsum_win = cumsum(ifelse(is.na(team_win), 0, team_win)),
        team_cumsum_loss = cumsum(ifelse(is.na(team_loss), 0, team_loss)),
        team_cumsum_game = team_cumsum_win + team_cumsum_loss,
        #Cumulative Rates
        team_cumrate_win = ifelse(team_cumsum_game == 0, NA, round(team_cumsum_win / team_cumsum_game, 3))) %>%
    ungroup()

all_team_data <- all_team_data %>%
    arrange(team_team, team_date) %>%
    group_by(team_team) %>%
    mutate(team_back2back = ifelse(difftime(team_date, lag(team_date), units = "days") == 1, 1, 0)) %>%
    replace_na(list(team_back2back = 0)) %>%
    ungroup()
```

------------------------------------------------------------------------

# Logistic Regression & Linear Regression

## Outcome 1: Margin of Victory: Binary Logistic Regression

### Pre-Analysis Cleaning

```{r}
in_wins <- all_team_data %>%
    filter(team_plusminus > 0) %>%
    group_by(team_team) %>%
    mutate(team_cummean_plusminus_win = round(cummean(team_plusminus))) %>%
    dplyr::select(team_team, team_date, team_cummean_plusminus_win)

in_losses <- all_team_data %>%
    filter(team_plusminus < 0) %>%
    group_by(team_team) %>%
    mutate(team_cummean_plusminus_loss = round(cummean(team_plusminus))) %>%
    dplyr::select(team_team, team_date, team_cummean_plusminus_loss)

all_team_data <- all_team_data %>%
    left_join(in_wins, by = c("team_team", "team_date")) %>%
    left_join(in_losses, by = c("team_team", "team_date"))

all_team_data <- all_team_data %>%
    group_by(team_team) %>%
    fill(team_cummean_plusminus_win, team_cummean_plusminus_loss, .direction = "down") %>%
    ungroup()

opponent_margins <- all_team_data %>%
    select(team_date, team_team, team_opponent, 
           team_cummean_plusminus_win, team_cummean_plusminus_loss) %>%
    rename("team_1" = "team_team", "team_2" = "team_opponent", 
           "opponent_cummean_plusminus_win"="team_cummean_plusminus_win", 
           "opponent_cummean_plusminus_loss"="team_cummean_plusminus_loss") %>%
    dplyr::select(-team_1)

all_team_data <- all_team_data %>%
    left_join(opponent_margins, by = c("team_team"="team_2", "team_date"))
```


```{r}
outcome_margin_model <- glm(team_game_outcome_1 ~ team_cummean_plusminus_win 
                            + opponent_cummean_plusminus_loss
                            , data = all_team_data, family = "binomial")
outcome_margin_model_summary <- summary(outcome_margin_model)
outcome_margin_model_summary

outcome_margin_2_model <- glm(team_game_outcome_1 ~ team_cummean_plusminus_win 
                            + opponent_cummean_plusminus_loss
                            + factor(team_team)
                            + factor(team_opponent)
                            , data = all_team_data, family = "binomial")
outcome_margin_2_model_summary <- summary(outcome_margin_2_model)
outcome_margin_2_model_summary
```

### TGO 1: Model Predictors Data Frame

```{r}
#Model Estimate to Main Data Frame
outcome_margin_fitted <- fitted(outcome_margin_model)
outcome_margin_predicted_values <- predict(outcome_margin_model, newdata = all_team_data, type = "response")
all_team_data <- cbind(all_team_data, outcome_margin_predicted_values)

#Cleaning Up Probability
all_team_data <- all_team_data %>%
    mutate(outcome_margin_predicted_values = round(100 * outcome_margin_predicted_values, 1))

#Model Estimate to Main Data Frame
outcome_margin_2_fitted <- fitted(outcome_margin_2_model)
outcome_margin_2_predicted_values <- predict(outcome_margin_2_model, newdata = all_team_data, type = "response")
all_team_data <- cbind(all_team_data, outcome_margin_2_predicted_values)

#Cleaning Up Probability
all_team_data <- all_team_data %>%
    mutate(outcome_margin_2_predicted_values = round(100 * outcome_margin_2_predicted_values, 1))
```


## Outcome 2: Game Outcome: Binary Logistic Regression

```{r}
winloss_model <- glm(team_game_outcome_2 ~ team_home
                 + team_cummean_offrating
                 + team_cummean_defrating
                 + opponent_cummean_offrating
                 + opponent_cummean_defrating
                 , data = all_team_data, family = "binomial")

winloss_model_summary <- summary(winloss_model)
winloss_model_summary

winloss_model_2 <- glm(team_game_outcome_2 ~ team_home
                 + team_cummean_offrating
                 + team_cummean_defrating
                 + opponent_cummean_offrating
                 + opponent_cummean_defrating
                 + factor(team_team)
                 + factor(team_opponent)
                 , data = all_team_data, family = "binomial")

winloss_model_2_summary <- summary(winloss_model_2)
winloss_model_2_summary
```

### TGO 2: Model Predictors Data Frame

```{r}
#Model Estimate to Main Data Frame
winloss_predictions <- fitted(winloss_model)
winloss_predicted_values <- predict(winloss_model, newdata = all_team_data, type = "response")
all_team_data <- cbind(all_team_data, winloss_predicted_values)

#Cleaning Up Probability
all_team_data <- all_team_data %>%
    mutate(winloss_predicted_values = round(100 * winloss_predicted_values, 1))

#Model Estimate to Main Data Frame
winloss_2_predictions <- fitted(winloss_model_2)
winloss_2_predicted_values <- predict(winloss_model_2, newdata = all_team_data, type = "response")
all_team_data <- cbind(all_team_data, winloss_2_predicted_values)
#Cleaning Up Probability
all_team_data <- all_team_data %>%
    mutate(winloss_2_predicted_values = round(100 * winloss_2_predicted_values, 1))
```

## Outcome 3: Game Outcome (w/ Previous Condition Met): Binary Logistic Regression

### Pre-Analysis Cleaning

```{r}
team_temporary_df_2 <- all_team_data %>%
    dplyr::select(team_date, team_team, team_opponent,
                  winloss_predicted_values, winloss_2_predicted_values) %>%
    rename("team_1" = "team_team", "team_2" = "team_opponent",
           "opponent_winloss_predicted_values"="winloss_predicted_values",
           "opponent_winloss_2_predicted_values"="winloss_2_predicted_values") %>%
    dplyr::select(-team_1)

all_team_data <- all_team_data %>%
    left_join(team_temporary_df_2, by = c("team_team"="team_2", "team_date"))

all_team_data <- all_team_data %>%
    mutate(winloss_model_guess = ifelse(winloss_predicted_values > opponent_winloss_predicted_values, 1, 0)) %>%
    mutate(winloss_model_2_guess = ifelse(winloss_2_predicted_values > opponent_winloss_2_predicted_values, 1, 0)) %>%
    #Outcome 3: Predicting Upset
    mutate(team_game_outcome_3a = case_when(
        winloss_model_guess == 1 & team_plusminus < 0 ~ 1,
        winloss_model_guess == 1 & team_plusminus > 0 ~ 0,
        TRUE ~ NA)) %>%
    mutate(team_game_outcome_3b = case_when(
        winloss_model_2_guess == 1 & team_plusminus < 0 ~ 1,
        winloss_model_2_guess == 1 & team_plusminus > 0 ~ 0,
        TRUE ~ NA))

all_team_data <- all_team_data %>%
    dplyr::group_by(team_team) %>%
    mutate(
    #Model 1
    team_cumsum_favored = cumsum(ifelse(is.na(winloss_model_guess), 0, winloss_model_guess)),
    team_cumsum_upset = cumsum(ifelse(is.na(team_game_outcome_3a), 0, team_game_outcome_3a)),
    #Model 2
    team_cumsum_favored_2 = cumsum(ifelse(is.na(winloss_model_2_guess), 0, winloss_model_2_guess)),
    team_cumsum_upset_2 = cumsum(ifelse(is.na(team_game_outcome_3b), 0, team_game_outcome_3b)),
    
    team_cumrate_favored = ifelse(team_cumsum_game == 0, NA, round(team_cumsum_favored / team_cumsum_game, 3)),
    team_cumrate_upset = ifelse(team_cumsum_favored == 0, NA, round(team_cumsum_upset / team_cumsum_favored, 3)),
    
    team_cumrate_favored_2 = ifelse(team_cumsum_game == 0, NA, round(team_cumsum_favored_2 / team_cumsum_game, 3)),
    team_cumrate_upset_2 = ifelse(team_cumsum_favored_2 == 0, NA, round(team_cumsum_upset_2 / team_cumsum_favored_2, 3))
    ) %>%
    ungroup()
```

### Model Analysis

```{r}
upset_model <- glm(team_game_outcome_3a ~ team_cumrate_upset
                   + opponent_cummean_offrating
                   + opponent_cummean_defrating
                   + team_back2back
                   , data = all_team_data, family = "binomial")
upset_model_summary <- summary(upset_model)
upset_model_summary
```

### TGO 3: Model Predictors Data Frame

```{r}
#Model Estimate to Main Data Frame
upset_predictions <- fitted(upset_model)
upset_predicted_values <- predict(upset_model, newdata = all_team_data, type = "response")
all_team_data <- cbind(all_team_data, upset_predicted_values)

#Cleaning Up Probability
all_team_data <- all_team_data %>%
    mutate(upset_predicted_values = round(100 * upset_predicted_values, 1)) %>%
    mutate(upset_predicted_values = case_when(
        winloss_predicted_values > opponent_winloss_predicted_values ~ upset_predicted_values,
        winloss_predicted_values < opponent_winloss_predicted_values ~ NA))
```

## Outcome 5: Game Total Over/Under: Linear Regression

```{r}
gametotal_model <- lm(game_points ~ team_cummean_pts 
                      + opponent_cummean_pts
                      + team_cummean_pts_allowed
                      + opponent_cummean_pts_allowed, data = all_team_data)
gametotal_model_summary <- summary(gametotal_model)
gametotal_model_summary
```

### TGO 5: Model Predictors to Data Frame

```{r}
gametotal_predicted_values <- predict(gametotal_model, newdata = all_team_data)
all_team_data <- cbind(all_team_data, gametotal_predicted_values)

all_team_data <- all_team_data %>%
    mutate(gametotal_predicted_values = round(gametotal_predicted_values))
```

# Data Export

```{r}
source("/Users/camsmithers/Desktop/NBA Project/Main/Code/Daily Data.R")

saveRDS(all_team_data, "/Users/camsmithers/Desktop/NBA Project/Main/Data/cleaned_nba_data.rds")

source("/Users/camsmithers/Desktop/NBA Project/Main/Code/Model Visualizations.R")
```
