---
title: "New Player Analysis"
author: "Cam Smithers"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sports Betting Analysis 2: Player Outcomes and Predictions

## Loading Packages

```{r}
library(readxl)
library(dplyr)
library(purrr)
library(ggplot2)
library(tidyverse)
library(readr)
library(purrr)
library(stringr)
library(broom)
library(parameters)
library(ggplot2)
```

# Data Cleaning

## Loading in All Data

```{r warning=FALSE}
#source("/Users/camsmithers/Desktop/NBA Project/Main/Data/DataScript.R")

all_player_data <- player_traditional %>%
    left_join(player_advanced, by = c("team", "date", "player")) %>% 
    left_join(player_misc, by = c("team", "date", "player")) %>%
    left_join(player_scoring, by = c("team", "date", "player")) %>%
    full_join(player_outcome, by = c("team", "date", "player")) #Used full join since data from traditional data doesn't exists yet
```

## Data Cleaning 1: General

```{r}
all_player_data <- all_player_data %>%
    select(-home, -pf.y, -blk.y, -home.y.y, -home.x.x, -home.y) %>%
    rename("home" = "home.x", "blk" = "blk.x", "pf" = "pf.x") %>%
    rename_with(~ paste0("player_", .)) %>%
    filter(!is.na(player_player))
```

## Data Cleaning 2: Important Outcomes

```{r}
cumsum_with_na <- function(variable) {
    ifelse(is.na(variable), NA, cumsum(ifelse(is.na(variable), 0, variable)))
}

all_player_data <- all_player_data %>%
  mutate(
    # Binary indicators and availability flags
    player_cover_pts = ifelse(player_pts > player_gen_player_pts, 1, 0),
    player_cover_reb = ifelse(player_reb > player_gen_player_reb, 1, 0),
    player_cover_ast = ifelse(player_ast > player_gen_player_ast, 1, 0),
    player_pts_odds_avi = ifelse(!is.na(player_gen_player_pts), 1, 0),
    player_reb_odds_avi = ifelse(!is.na(player_gen_player_reb), 1, 0),
    player_ast_odds_avi = ifelse(!is.na(player_gen_player_ast), 1, 0)
  ) %>%
  group_by(player_player) %>%
  mutate(
    # General sums
    player_pts_odds_sum = sum(player_pts_odds_avi, na.rm = TRUE),
    player_reb_odds_sum = sum(player_reb_odds_avi, na.rm = TRUE),
    player_ast_odds_sum = sum(player_ast_odds_avi, na.rm = TRUE),
    player_cover_pts_sum = sum(player_cover_pts, na.rm = TRUE),
    player_cover_reb_sum = sum(player_cover_reb, na.rm = TRUE),
    player_cover_ast_sum = sum(player_cover_ast, na.rm = TRUE),
    
    # Cumulative sums with NA handling
    player_pts_odds_cumsum = cumsum(ifelse(is.na(player_pts_odds_avi), 0, player_pts_odds_avi)),
    player_reb_odds_cumsum = cumsum(ifelse(is.na(player_reb_odds_avi), 0, player_reb_odds_avi)),
    player_ast_odds_cumsum = cumsum(ifelse(is.na(player_ast_odds_avi), 0, player_ast_odds_avi)),
    player_cover_pts_cumsum = cumsum(ifelse(is.na(player_cover_pts), 0, player_cover_pts)),
    player_cover_reb_cumsum = cumsum(ifelse(is.na(player_cover_reb), 0, player_cover_reb)),
    player_cover_ast_cumsum = cumsum(ifelse(is.na(player_cover_ast), 0, player_cover_ast)),
    
    # General rates (avoiding division by zero or NA)
    player_cover_pts_gen_rate = ifelse(player_pts_odds_sum == 0, NA, round(player_cover_pts_sum / player_pts_odds_sum, 2)),
    player_cover_reb_gen_rate = ifelse(player_reb_odds_sum == 0, NA, round(player_cover_reb_sum / player_reb_odds_sum, 2)),
    player_cover_ast_gen_rate = ifelse(player_ast_odds_sum == 0, NA, round(player_cover_ast_sum / player_ast_odds_sum, 2)),
    
    # Cumulative rates (avoiding division by zero or NA)
    player_cover_pts_cumsum_rate = ifelse(player_pts_odds_cumsum == 0, NA, round(player_cover_pts_cumsum / player_pts_odds_cumsum, 2)),
    player_cover_reb_cumsum_rate = ifelse(player_reb_odds_cumsum == 0, NA, round(player_cover_reb_cumsum / player_reb_odds_cumsum, 2)),
    player_cover_ast_cumsum_rate = ifelse(player_ast_odds_cumsum == 0, NA, round(player_cover_ast_cumsum / player_ast_odds_cumsum, 2))
  ) %>%
  ungroup()

```

## Data Cleaning 3: Player Averages

```{r}
average_player <- all_player_data %>%
    group_by(player_player) %>%
    summarize(
        player_pts = round(mean(player_pts, na.rm = TRUE), 1), 
        player_reb = round(mean(player_reb, na.rm = TRUE), 1), 
        player_ast = round(mean(player_ast, na.rm = TRUE), 1), 
        player_tspct = round(mean(player_tspct, na.rm = TRUE), 1),
        player_offrating = round(mean(player_offrating, na.rm = TRUE), 1), 
        player_defrating = round(mean(player_defrating, na.rm = TRUE), 1), 
        player_tov = round(mean(player_tov, na.rm = TRUE), 1)
    ) %>%
    rename_with(~ paste0("avg_", .))

average_player_last10 <- all_player_data %>%
    group_by(player_player) %>%
    slice_tail(n = 10) %>%
    summarize(
        player_pts = round(mean(player_pts, na.rm = TRUE), 1), 
        player_reb = round(mean(player_reb, na.rm = TRUE), 1), 
        player_ast = round(mean(player_ast, na.rm = TRUE), 1), 
        player_tspct = round(mean(player_tspct, na.rm = TRUE), 1),
        player_offrating = round(mean(player_offrating, na.rm = TRUE), 1), 
        player_defrating = round(mean(player_defrating, na.rm = TRUE), 1), 
        player_tov = round(mean(player_tov, na.rm = TRUE), 1)
    ) %>%
    rename_with(~ paste0("avg_", ., "_last10"))

average_player_full <- average_player %>%
    left_join(average_player_last10, by = c("avg_player_player" = "avg_player_player_last10")) %>%
    mutate(player_weighted_pts  = (avg_player_pts * .50) + (avg_player_pts_last10 * .50)) %>%
    mutate(player_weighted_reb  = (avg_player_reb * .50) + (avg_player_ast_last10 * .50)) %>%
    mutate(player_weighted_ast  = (avg_player_ast * .50) + (avg_player_reb_last10 * .50)) %>%
    mutate(player_weighted_tspct  = (avg_player_tspct * .50) + (avg_player_tspct_last10 * .50)) %>%
    mutate(player_weighted_offrating  = (avg_player_offrating * .50) + (avg_player_offrating_last10 * .50)) %>%
    mutate(player_weighted_defrating  = (avg_player_defrating * .50) + (avg_player_defrating_last10 * .50)) %>%
    mutate(player_weighted_tov  = (avg_player_tov * .50) + (avg_player_tov_last10 * .50))
    
all_player_data <- all_player_data %>%
    left_join(average_player_full, by = c("player_player" = "avg_player_player"))
```

## Data Cleaning 4: Player & Team Data

```{r}
all_data <- all_player_data %>%
    left_join(all_team_data, by = c("player_team" = "team_team", "player_date" = "team_date"))
```

# Modeling & Data Cleaning

## Model Testing 1: Player Points

```{r}
player_pts_model <- glm(player_cover_pts ~ avg_player_pts
                        + player_gen_player_pts
                        + avg_opponent_team_defrating
                        , data =  all_data, family = binomial)
summary_player_pts_model <- summary(player_pts_model)
summary_player_pts_model

player_pts_model_v2 <- glm(player_cover_pts ~ avg_player_pts
                        + player_gen_player_pts
                        + avg_opponent_team_defrating
                        + player_cover_pts_gen_rate
                        , data =  all_data, family = binomial)
summary_player_pts_model_v2 <- summary(player_pts_model_v2)
summary_player_pts_model_v2

player_pts_model_v3 <- glm(player_cover_pts ~ avg_player_pts
                           + player_gen_player_pts
                           + avg_opponent_team_defrating
                           + player_cover_pts_cumsum_rate
                           , data = all_data, family = binomial)
summary_player_pts_model_v3 <- summary(player_pts_model_v3)
summary_player_pts_model_v3
```

### Model Data Creating 1: Cover Player Points

```{r}
player_pts_df <- as.data.frame(summary_player_pts_model$coefficients) %>%
    rownames_to_column(var = "Variable") %>%
    as_tibble() %>%
    filter(`Pr(>|z|)` < .05)

player_pts_df_v2 <- as.data.frame(summary_player_pts_model_v2$coefficients) %>%
    rownames_to_column(var = "Variable") %>%
    as_tibble() %>%
    filter(`Pr(>|z|)` < .05)

player_pts_df_v3 <- as.data.frame(summary_player_pts_model_v3$coefficients) %>%
    rownames_to_column(var = "Variable") %>%
    as_tibble() %>%
    filter(`Pr(>|z|)` < .05)
```

### Data Cleaning 5: Adding Player Points Model

```{r}
all_data <- all_data %>%
    mutate(cover_pts_prob = round(100 * (1 / (1 + e^-((player_pts_df$Estimate[1]) 
           + (player_pts_df$Estimate[2] * avg_player_pts) 
           + (player_pts_df$Estimate[3] * player_gen_player_pts) 
           + (player_pts_df$Estimate[4] * avg_opponent_team_defrating)))), 1)
           ) %>%
    mutate(cover_pts_v2_prob = round(100 * (1 / (1 + e^-((player_pts_df_v2$Estimate[1]) 
           + (player_pts_df_v2$Estimate[2] * avg_player_pts) 
           + (player_pts_df_v2$Estimate[3] * player_gen_player_pts) 
           + (player_pts_df_v2$Estimate[4] * avg_opponent_team_defrating)
           + (player_pts_df_v2$Estimate[5] * player_cover_pts_gen_rate)))), 1)
           ) %>%
    mutate(cover_pts_v3_prob = round(100 * (1 / (1 + e^-((player_pts_df_v3$Estimate[1]) 
           + (player_pts_df_v3$Estimate[2] * avg_player_pts) 
           + (player_pts_df_v3$Estimate[3] * player_gen_player_pts) 
           + (player_pts_df_v3$Estimate[4] * avg_opponent_team_defrating)
           + (player_pts_df_v3$Estimate[5] * player_cover_pts_cumsum_rate)))), 1)
           )
```

## Model Testing 2: Player Rebounds

```{r}
player_reb_model <- glm(player_cover_reb ~ player_gen_player_reb
                        + player_cover_reb_cumsum_rate
                        , data = all_data, family = binomial)
summary(player_reb_model)
```


## Model Testing 3: Player Assists

